---
title: "Geographic Patterns of Eviction"
subtitle: "Where evictions concentrate in Tulsa County"
format:
  okpolicy-website-template-html: default
title-block-banner: "#2f507f"
title-block-banner-color: "white"
fontcolor: "black"
linkcolor: "black"
number-sections: true
smooth-scroll: true
knitr:
  opts_chunk:
    fig.align: center
    echo: false
    message: false
    warning: false
description: |
  Spatial analysis of granted eviction filings in Tulsa County using
  geocoded court records. Identifies geographic concentration areas
  and examines how eviction density varies across the county.
---

```{r setup}
#| label: setup
#| include: false

# --- Pipeline output consumption ---
# Same pattern as the trends report: read from pipeline output.
# Currently reads raw CSV; will switch to parquet in Phase 3.

library(dplyr)
library(ggplot2)
library(gt)
library(arrow)

evictions <- readr::read_csv(
  here::here("data", "raw", "evictions_granted_tulsa_2021_to_2025.csv"),
  show_col_types = FALSE
)
```

## What This Report Shows

Evictions don't happen uniformly across a city. This report uses geocoded
court records to visualize **where** granted evictions concentrate in Tulsa
County. Each data point represents a court case where eviction was granted
and the filing address was geocoded with high confidence (accuracy ≥ 0.85).

The goal is not to explain *why* evictions cluster — that requires
socioeconomic context this dataset doesn't contain. The goal is to surface
*where* they cluster so that further investigation can focus on the right
areas.

::: {.content-hidden when-profile="public"}
### Draft Note

A proper spatial analysis would use `sf` and map tiles. For the learning
project, we use `ggplot2` density plots with lat/lon coordinates. This
demonstrates the pipeline → report pattern without adding GIS complexity.
:::

## Spatial Distribution

```{r density-map}
#| label: fig-density-map
#| fig-cap: "Density of granted evictions across Tulsa County"
#| fig-width: 10
#| fig-height: 8

# Filter to reasonable Tulsa metro bounding box
tulsa_evictions <- evictions |>
  filter(
    lat > 35.9, lat < 36.4,
    lon > -96.2, lon < -95.7
  )

ggplot(tulsa_evictions, aes(x = lon, y = lat)) +
  stat_density_2d(
    aes(fill = after_stat(level)),
    geom = "polygon",
    alpha = 0.6
  ) +
  geom_point(alpha = 0.03, size = 0.3, color = "#333333") +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  coord_fixed(ratio = 1.3) +
  labs(
    title = "Eviction Density in Tulsa County",
    subtitle = "Granted evictions, 2021–2025 | Each point = one case",
    x = "Longitude",
    y = "Latitude",
    fill = "Density",
    caption = "Source: Oklahoma court records via ojodb | Geocoded, accuracy ≥ 0.85"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold"),
    plot.title.position = "plot",
    panel.grid = element_line(color = "#eeeeee")
  )
```

## Concentration by Year

```{r concentration-by-year}
#| label: fig-year-facets
#| fig-cap: "Eviction density by year — has the pattern shifted?"
#| fig-width: 12
#| fig-height: 10

ggplot(tulsa_evictions, aes(x = lon, y = lat)) +
  stat_density_2d(
    aes(fill = after_stat(level)),
    geom = "polygon",
    alpha = 0.6
  ) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  coord_fixed(ratio = 1.3) +
  facet_wrap(~filing_year, ncol = 3) +
  labs(
    title = "Eviction Density by Year",
    subtitle = "Tulsa County, 2021–2025",
    x = "Longitude",
    y = "Latitude",
    fill = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.title.position = "plot",
    strip.text = element_text(face = "bold", size = 13),
    panel.grid = element_line(color = "#eeeeee")
  )
```

## Geocoding Quality

```{r geocoding-quality}
#| label: tbl-geocoding

evictions |>
  mutate(
    accuracy_band = case_when(
      geo_accuracy == 1 ~ "Exact (1.0)",
      geo_accuracy >= 0.95 ~ "High (0.95–0.99)",
      geo_accuracy >= 0.90 ~ "Good (0.90–0.94)",
      geo_accuracy >= 0.85 ~ "Acceptable (0.85–0.89)",
      TRUE ~ "Below threshold"
    )
  ) |>
  count(accuracy_band, name = "n_cases") |>
  arrange(desc(n_cases)) |>
  mutate(pct = scales::percent(n_cases / sum(n_cases), accuracy = 0.1)) |>
  gt::gt() |>
  gt::cols_label(
    accuracy_band = "Accuracy Band",
    n_cases = "Cases",
    pct = "Share"
  ) |>
  gt::fmt_number(columns = n_cases, decimals = 0) |>
  gt::tab_header(
    title = "Geocoding Accuracy Distribution",
    subtitle = "All granted evictions in dataset"
  )
```

## Key Observations

The density maps show whether eviction filings cluster in specific
neighborhoods or distribute broadly. When clusters are stable across years,
they often correspond to areas with high concentrations of rental housing,
lower-income demographics, or specific property management patterns.

The geocoding quality table confirms that the vast majority of records have
exact or near-exact location data, meaning the spatial patterns are reliable
rather than artifacts of geocoding errors.

---

*Data pipeline: `targets::tar_make()` → `data/parquet/` → this report.*
