---
title: "Eviction Trends in Tulsa County"
subtitle: "Year-over-year filing patterns, 2021–2025"
format:
  okpolicy-website-template-html: default
title-block-banner: "#2f507f"
title-block-banner-color: "white"
fontcolor: "black"
linkcolor: "black"
number-sections: true
smooth-scroll: true
knitr:
  opts_chunk:
    fig.align: center
    echo: false
    message: false
    warning: false
description: |
  How has the volume of granted evictions in Tulsa changed from 2021 to 2025?
  This report examines year-over-year trends using geocoded court records
  processed through the project's targets pipeline.
---

```{r setup}
#| label: setup
#| include: false

# --- Pipeline output consumption (Phase 3) ---
# This report now reads from parquet files produced by the targets pipeline.
# The pipeline writes summarized data to data/parquet/ which this report
# consumes. This is the core pattern: pipeline produces → report consumes.
#
# If the parquet file doesn't exist yet (pipeline hasn't been run),
# fall back to reading the raw CSV directly.

library(dplyr)
library(ggplot2)
library(gt)
library(arrow)

# Try parquet first (pipeline output), fall back to raw CSV
parquet_path <- here::here("data", "parquet", "eviction_summary.parquet")
raw_path <- here::here("data", "raw", "evictions_granted_tulsa_2021_to_2025.csv")

if (file.exists(parquet_path)) {
  annual_counts <- arrow::read_parquet(parquet_path)
  data_source <- "parquet (pipeline output)"
} else {
  evictions_raw <- readr::read_csv(raw_path, show_col_types = FALSE)
  annual_counts <- evictions_raw |>
    count(filing_year, name = "n") |>
    mutate(
      change = n - lag(n),
      pct_change = change / lag(n)
    )
  data_source <- "raw CSV (pipeline not yet run)"
}
```

## What This Report Shows

This report tracks the volume of **granted eviction filings** in Tulsa County
from 2021 through 2025. "Granted" means a court issued a judgment in favor of
eviction — not every filing results in one. The data comes from Oklahoma court
records, geocoded and filtered to cases with high-confidence location data
(accuracy ≥ 0.85).

::: {.content-hidden when-profile="public"}
### Data Source
Currently reading from: **`r data_source`**.
Run `targets::tar_make()` to generate pipeline outputs.
:::

## Annual Volume

```{r annual-volume}
#| label: fig-annual-volume
#| fig-cap: "Granted evictions per year in Tulsa County"

ggplot(annual_counts, aes(x = filing_year, y = n)) +
  geom_col(fill = "#2f507f", width = 0.7) +
  geom_text(
    aes(label = scales::comma(n)),
    vjust = -0.5, size = 4, fontface = "bold"
  ) +
  scale_x_continuous(breaks = annual_counts$filing_year) +
  scale_y_continuous(
    labels = scales::comma,
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(
    title = "Granted Evictions by Year",
    subtitle = "Tulsa County, 2021–2025",
    x = "Filing Year", y = "Number of Granted Evictions",
    caption = "Source: Oklahoma court records via ojodb | Geocoded, accuracy ≥ 0.85"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold"),
    plot.title.position = "plot"
  )
```

## Summary Table

```{r summary-table}
#| label: tbl-annual-summary

annual_counts |>
  gt::gt() |>
  gt::cols_label(
    filing_year = "Year", n = "Evictions Granted",
    change = "Change", pct_change = "% Change"
  ) |>
  gt::fmt_number(columns = c(n, change), decimals = 0) |>
  gt::fmt_percent(columns = pct_change, decimals = 1) |>
  gt::sub_missing(missing_text = "—") |>
  gt::tab_header(
    title = "Annual Eviction Summary",
    subtitle = "Tulsa County, granted evictions with geocoded addresses"
  ) |>
  gt::tab_source_note("Source: Oklahoma court records | Geocoded accuracy ≥ 0.85")
```

## What the Data Suggests

The trend line tells a story about housing stability in Tulsa County. Changes
in eviction volume can reflect shifts in housing policy, economic conditions,
pandemic-era protections expiring, or changes in landlord behavior.

---

*Data pipeline: `targets::tar_make()` → `data/parquet/` → this report.*
